---
title: "lab_6_bg"
author: "Bridget Gibbons"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
#attach packages

library(tidyverse)
library(here)
library(raster)
library(janitor)
library(sf)
library(tmap)
library(tmaptools)
library(gstat)

```

## Grand Canyon GeoTIFF

```{r}
gc_dem <- raster(here("data", "gc_dem.tif"))

# Look at it using plot():

plot(gc_dem)

# Check the CRS:
gc_dem@crs

# Check the extent (bounding box):
gc_dem@extent

# Run the CRS code again (up above), and then copy the arguments starting with the plus sign to the end
# Create a wgs84 with latitude and longitude metrics instead (it's currently in utm)
# Paste, change the proj, delete the stuff at the end
# or could put in "proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

wgs84 <- "+proj=longlat +zone=12 +datum=WGS84 +units=m +no_defs" 

gc_reproj <- projectRaster(gc_dem, crs = wgs84, method = "bilinear")

gc_reproj@extent

```


### Crop raster to a smaller area (bounding box):

```{r}

bounds <- as(extent(-112.4, -112.0, 36.1, 36.3), 'SpatialPolygons')

```

### Make the CRS of our bounding boz the same as for gc_reproj:

```{r}
crs(bounds) <- crs(gc_reproj)
```

### Now let's crop our original spatial data:
```{r}

gc_crop <- crop(gc_reproj, bounds)

plot(gc_crop)

```

### Resample using the `aggregate()` function:

```{r}
gc_agg <- raster::aggregate(gc_crop, fact = 10)

plot(gc_agg)
```

### Now let's get safe & sound in ggplot:

First, convert data to a data frame:

```{r}
gc_df <- as.data.frame(gc_agg, xy = TRUE)

ggplot(data = gc_df, aes(x=x, y=y)) +
  geom_raster(aes(fill = gc_dem)) +
  coord_quickmap() +
  theme_minimal() +
  scale_fill_gradientn(colors = c("purple",
                                  "magenta",
                                  "orange",
                                  "yellow", 
                                  "white")
                       )
  
```

### How can I just select cells that match given criteria?

```{r}

gc_hab <- gc_crop

# Set any cells outside of 1000 - 1500 to NA
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

# Plot.
plot(gc_hab)

# Cool! 

```

Now let's make this interactive with tmap:

```{r}
tmap_mode("view")

tm_shape(gc_hab) +
  tm_raster(legend.show = FALSE, palette = "plasma")

# try getting tmap directly from devtools via github?
```

### Kriging rain in Kansas

Read in the KS counties shapefile data

```{r}
ks_counties <- read_sf(here("data", "ks_counties", "ks_counties_shapefile.shp"))

# Use base plot:
plot(ks_counties)

#Check CRS:
st_crs(ks_counties) #there isn't one, so we should set one

# Set to EPSG 4326:
st_crs(ks_counties) <- 4326

plot(ks_counties)

```

Practice in ggplot:
```{r}

ggplot(data = ks_counties) +
  geom_sf()

```

Now, read in the rainfall data:

```{r}
ks_rain <- read_csv(here("data", "ks_rain.csv")) %>% 
  clean_names() # R doesn't know that this is spatial data yet, just knows it has a column called lat and lon
```

Update the ks_rain data to be recognized as spatial points:

```{r}

ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs = 4326) # in the order of longitude and then latitude always, not the order of the columns

```

```{r}
ggplot() +
  geom_sf(data = ks_counties) +
  geom_sf(data = ks_sf,
          aes(color = amt, size = amt))
```

## Kriging to predict rainfall:

```{r}
ks_sp <- as_Spatial(ks_sf)
class(ks_sp)
```

### Make a spatial pixels grid that we'll make predictions over:

```{r}
#can look at bbox(ks_sp) to see the range of points we need to include

lat <- seq(37, 40, length.out = 200) #creates a sequence of values from 37.0 to 40.0 with 200 evenly spaced values between them
long <- seq(-94.6, -102, length.out = 200)

# Now we're going to make this into a spatial grid:
grid <- expand.grid(lon = long, lat = lat)
grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)


plot(grid_sp)
```

### Now, let's make a variogram (from the rainfall data)

```{r}
ks_vgm <- variogram(amt ~ 1, data = ks_sp)
plot(ks_vgm)

# estimate variogram parameters
# guess nugget is around 0.1
# guess sill is around 1.0
# guess range is around 200

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.1, psill = 1.0, range = 200, model = "Sph"))

plot(ks_vgm, ks_vgm_fit)

```

### Now... krige!

```{r}

ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)

```

```{r}
spplot(ks_krige, "var1.pred")
```

### Make a data frame of kriged predictions

```{r}
ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1,
         latitude = coords.x2)

# Convert this data frame to sf object:
rain_sf <- st_as_sf(ks_df, coords = c("longitude", "latitude"), crs = 4326)

ggplot(rain_sf) +
  geom_sf(aes(color = var1.pred))
```

### Crop to the actual Kansas outline:

```{r}
ks <- read_sf(here("data", "states"),
              layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)
```

```{r}
# Find thee intersection of the two:
rain_sf_ks <- st_intersection(rain_sf, ks)

ggplot(data = rain_sf_ks) +
  geom_sf(aes(color = var1.pred))
```

